---
title: "DS2_HW3"
author: "Yixiao Sun"
date: "2024-03-22"
output: pdf_document
---

```{r}
library(dplyr)
library(caret)
library(glmnet) 
library(tidymodels) 
library(pROC) 
library(MASS)
library(caTools)
data <- read.csv("~/Desktop/P8106 Data Science 2/ds2_hw3/auto.csv")
set.seed(123)
data <- data %>% mutate(mpg_cat = as.factor(mpg_cat), origin = as.factor(origin))
contrasts(data$mpg_cat)
ctrl1 <- trainControl(method = "cv", number = 10,summaryFunction = twoClassSummary,classProbs = TRUE)

data_split <- initial_split(data, prop = 0.7) 
train_data <- training(data_split) 
test_data <- testing(data_split)

```

# Question 1
```{r}
set.seed(123)
enet.fit <- train(mpg_cat ~ ., 
                   data = train_data, 
                   method = "glmnet",
                   tuneGrid = expand.grid(alpha = seq(0,1,length = 21),
                                          lambda = exp(seq(1, -8, length = 100))),
                  metric = "ROC",
                  trControl = ctrl1) 
enet.fit$bestTune
print(coef(enet.fit$finalModel,enet.fit$bestTune$lambda))
plot(enet.fit, xTrans = log)
```


# Based on the enet model which can be helpful to identify the redundant variable, we identify variables cylinders and displacement as redundant variable.


# Question 2
```{r}
set.seed(123)
predict_prob <- predict(enet.fit, newdata = test_data, type = "prob")[,2]
predicted_class <- ifelse(predict_prob > 0.5, "low", "high")
CM<-confusionMatrix(data = as.factor(predicted_class), reference = test_data$mpg_cat, positive = "low")
```

# For the confusion matrix, we obtain an Accuracy of 0.8983, which indicates that our Elastic Net Model has an identify accuracy of 0.8983 .And we obtain a sensitivity of 0.9167, indicating that our Elatic Net Model classifies 91.67% of true “high” instances correctly. We get a specificity of 0.8793, indicating that our Elastic Net Model classifies 87.93% of true “low” instances correctly. Finally we get a Kappa of 0.7964, indicating that our model has a good performance in this classification.

# Question 3
```{r, warning = FALSE}
set.seed(123)
mars.fit <- train(mpg_cat ~.,
                    data = train_data,
                    method = "earth",
                    tuneGrid = expand.grid(degree = 1:4,
                                           nprune = 2:20),
                  metric = "ROC",
                  trControl = ctrl1)
plot(mars.fit)

bwplot(resamples(list(enet = enet.fit, mars = mars.fit)), matrix = "ROC")
```

# After creating the mars model, we conclude that the Elastic Net Model is still better for the prediction performance for producing a higher ROC value.

# Question 4
```{r}
set.seed(123)
library(MASS)
lda.fit <- train(mpg_cat ~ ., 
                   data = train_data,
                   method = "lda",
                   metric = "ROC",
                   trControl = ctrl1)
lda<-lda(mpg_cat ~ ., data = train_data)
plot(lda)
```

# Question 5
```{r}
bwplot(resamples(list(enet = enet.fit, mars = mars.fit, lda = lda.fit)), matrix = "RMSE")

resamp<-resamples(list(enet = enet.fit, mars = mars.fit, lda = lda.fit))
summary(resamp)
library(pROC)
roc_response <- roc(response = test_data$mpg_cat, predictor = as.numeric(predict_prob))
plot(roc_response,print.auc = T, legacy.axes = T)
plot(smooth(roc_response),add = T)
auc(roc_response)
error_rate <- 1- CM$overall
print(error_rate)
```

# Based on the comparison, we choose Elastic Net Model to predict the response variable, The AUC value is 0.9626, and the calculated misclassification error rate is 0.10169492.
